Generalization Errors of Hard Parameter Sharing: High-dimensional Asymptotics and Implications

### 2020-09-13 arxiv/proof_main.tex
revise

%\subsection{The Bias and Variance Asymptotics}
\paragraph{Proof overview.} 
%\todo{merge} Finally, in this subsection, we describe briefly the main ideas in the proofs of Lemmas \ref{lem_cov_shift} and \ref{lem_cov_derivative}.
%We first describe the proof of Theorem \ref{lem_cov_shift_informal}.
%We briefly describe the key ideas for proving Lemma \ref{lem_cov_shift} and Lemma \ref{lem_cov_derivative}.
We use the standard Stieltjes transform method (or the resolvent method) in random matrix theory \cite{bai2009spectral,tao2012topics,erdos2017dynamical}. Both Lemma \ref{lem_cov_shift} and Lemma \ref{lem_cov_derivative} require a study of the matrix inverse $(X_1^\top X_1+X_2^\top X_2)^{-1}$ for $X_1= Z_1\Sigma_1^{1/2}$ and $X_2= Z_2\Sigma_2^{1/2}$.
%\HZ{Suggestion: focus on the proof overview of Lemma \ref{lem_cov_shift} only. Use one paragraph in the end to say how to extend the idea to Lemma \ref{lem_cov_derivative}}
Suppose $M=\Sig_1^{1/2} \Sig_2^{-1/2}$ has a singular value decomposition
\be\label{eigen2}
M= U\Lambda V^\top, \quad \Lambda=\text{diag}( \lambda_1, \ldots, \lambda_p).
\ee
Then using equation \eqref{eigen2}, we can write
\be\label{eigen2extra}(X_1^\top X_1+X_2^\top X_2)^{-1}= \Sigma_2^{-1/2}V\left(   \Lambda U^\top Z_1^\top Z_1 U\Lambda  + V^\top Z_2^\top Z_2V\right)^{-1}V^\top\Sigma_2^{-1/2}.\ee
We divide the discussion into the following parts.

@@

\noindent\textbf{Defining the resolvents.} For our purpose, we introduce a convenient \alert{self-adjoint linearization trick}\HZ{what does this trick mean? use less technical words},

@@

which has been proved to be useful in studying the local laws of random matrices of the Gram type \cite{Anisotropic, AEK_Gram, XYY_circular,DY20201}. We define the following $(p+n)\times (p+n)$ symmetric block matrix, which is a linear function of $Z_1$ and $Z_2$:

@@

    H(Z_1,Z_2): = n^{-1/2}\left( {\begin{array}{*{20}c}
   { 0 } & \Lambda U^{\top}Z_1^\top & V^\top Z_2^\top  \\
   {Z_1 U\Lambda  } & {0} & 0 \\
   {Z_2V} & 0 & 0

@@

For simplicity of notations, we define the index sets

@@

 We will consistently use the latin letters $i,j\in\sI_{0}$, greek letters $\mu,\nu\in\sI_{1}\cup \sI_{2}$, and $\fa,\fb,\mathfrak c\in \cal I$ \HZ{This $\fa, \fb$ notation is ugly. Why not just use $a,b$?}. Correspondingly, the indices of the matrices $Z_1$ and $Z_2$ are labelled as

@@

 Z_1= (z_{\mu i}:i\in \mathcal I_0, \mu \in \mathcal I_1), \quad Z_2= (z_{\nu i}:i\in \mathcal I_0, \nu \in \mathcal I_2).\ee

@@

 G (Z_1,Z_2,z):= \left[H(Z_1,Z_2)-\left( {\begin{array}{*{20}c}

@@

\end{array}} \right)\right]^{-1} , \quad z\in \mathbb C .

@@

and the resolvent of $  n^{-1}\left(\Lambda U^\top Z_1^\top Z_1 U\Lambda  + V^\top Z_2^\top Z_2V\right)$ as

@@

\cal G(z):=\left( n^{-1}  \Lambda U^\top Z_1^\top Z_1 U\Lambda  + n^{-1}V^\top Z_2^\top Z_2V -z\right)^{-1},\quad z\in \C.

@@

@@

With Schur complement formula, we can check that \HZ{Has $G(z)$ been defined before?}

@@

   { \cal G (z)} & \cal G(z)W  \\
   W^\top \cal G(z) & z (W^\top W - z)^{-1}

@@

 where we abbreviated $W:=n^{-1/2}(\Lambda U^\top Z_1^\top, V^\top Z_2^\top)$. \HZ{$W$ is overloaded with the output layers} This shows that a control of $G(z)$ yields directly a control of $\mathcal G(z)$ as the upper-left block. On the other hand, $G(z)$ is a little  easier to use than $\cal G(z)$, and obviously contains more information.

@@

\medskip
\noindent\textbf{Defining the asymptotic limits of the resolvents.}
%\label{sec aymp_limit_G}
%\subsection{Resolvents and limiting law}
We now describe the asymptotic limit of $G(z)$ as $n\to \infty$. First we define the deterministic limits of $(m_1(z), m_{2}(z))$ by $\left(-\frac{\rho_1+\rho_2}{\rho_1}a_{1}(z),-\frac{\rho_1+\rho_2}{\rho_2}a_{2}(z)\right)$, where $(a_1(z), a_2(z))$ is
%\HZ{This $M_{1}, M_{2}$ notation is awkward; consider changing it to ${M}_1(z), {M}_2(z)$ (or  better, say $a_1(z), a_2(z)$? if they correspond to $a_1, a_2$ when $z=0$)},
the unique solution to the following system of equations
\begin{equation}\label{selfomega_a}
\begin{split}
& \frac{\rho_1}{a_{1}(z)} = \frac{1}{p}\sum_{i=1}^p \frac{\lambda_i^2}{ - z+\lambda_i^2 a_{1}(z) +a_{2} (z) } + (\rho_1+\rho_2),\  \frac{\rho_2}{a_{2}(z)} = \frac{1}{p}\sum_{i=1}^p \frac{1 }{  -z+\lambda_i^2 a_{1}(z) +  a_{2}(z)  }+ (\rho_1+\rho_2) ,
\end{split}
\ee
satisfying that $\im a_{1}(z)>0$ and $\im a_{2}(z)>0$ for $z\in \C_+$ with $\im z$.
%Here, for simplicity of notations, we introduced the following ratios
%\be\label{ratios}
% \gamma_n :=\frac{p}{n}=\frac{1}{\rho_1+\rho_2},\quad r_1 :=\frac{n_1}{n}=\frac{\rho_1}{\rho_1+\rho_2},\quad r_2 :=\frac{n_2}{n}=\frac{\rho_2}{\rho_1+\rho_2}.
%\ee
We define the matrix limit of $G(z)$ as
\be \label{defn_piw}
\Gi(z) := \begin{pmatrix} (-z\id_p+a_{1}(z)\Lambda^2  +  a_{2}(z)\id_p)^{-1} & 0 & 0 \\ 0 & - \frac{\rho_1+\rho_2}{\rho_1} a_{1}(z)\id_{n_1} & 0 \\ 0 & 0 & -\frac{\rho_1+\rho_2}{\rho_2}a_{2}(z)\id_{n_2}  \end{pmatrix}.\ee
In particular, the matrix limit of $\cal G(z)$ is given by $(-z\id_p+a_{1}(z)\Lambda^2 + a_{2}(z)\id_p)^{-1}$.

\medskip
\noindent{\bf Deriving the bias and variance asymptotics.} Using definition \eqref{mainG}, we can write equation \eqref{eigen2extra} as
\be\label{rewrite X as R} (X_1^\top X_1+X_2^\top X_2)^{-1}=n^{-1}\Sigma_2^{-1/2}V\cal G(0)V^\top\Sigma_2^{-1/2}.\ee
When $z=0$, it is easy to check that equation \eqref{selfomega_a} can be reduced to \eqref{eq_a12extra}, which means that we actually have $a_1(0)=a_1$ and $a_2(0)=a_2$. Hence the matrix limit of $\cal G(0)$ is given by $(a_{1}\Lambda^2 + a_{2}\id_p)^{-1}$. \alert{Together with equation \eqref{rewrite X as R} and $V \Lambda^2 V^\top=M^\top M=\Sigma_2^{-1/2}\Sigma_1\Sigma_2^{-1/2}$ by equation \eqref{eigen2}, we can obtain equation \eqref{lem_cov_shift_eq} in the large $n$ limit.}\HZ{I do not understand this sentence.}
\noindent{\bf Gaussian case.} Now we give a heuristic derivation of the matrix limit when the entries of $Z_1$ and $Z_2$ are i.i.d. Gaussian. In this case,
%However, notice that if the entries of $Z_1\equiv Z_1^{\text{Gauss}}$ and $Z_2\equiv Z_2^{\text{Gauss}}$ are i.i.d. Gaussian, then
by the rotational invariance of the multivariate Gaussian distribution we have
\be\label{eq in Gauss} Z_1 U\Lambda \stackrel{d}{=} Z_1 \Lambda, \quad Z_2 V \stackrel{d}{=} Z_2,\ee
where ``$\stackrel{d}{=}$" means ``equal in distribution". Hence it suffices to consider the following resolvent
 \begin{equation} \label{resolv Gauss1}
   G(z)= \left( {\begin{array}{*{20}c}
   { -z\id_{p} } & n^{-1/2}\Lambda Z_1^\top & n^{-1/2} Z_2^\top  \\
   {n^{-1/2} Z_1 \Lambda  } & {-\id_{n_1}} & 0 \\
   {n^{-1/2} Z_2} & 0 & {-\id_{n_2}}
   \end{array}} \right)^{-1}.
\noindent{\bf Schur complements and concentration.}
%Now we discuss about how to obtain the matrix limit in equation \eqref{defn_piw}.
%\HZ{Divide into several parts like sec 7 to improve readability.}
The core quantities of the derivation are the following resolvent minors, which are defined by removing certain rows and columns of the matrix $H$.
\begin{definition}[Resolvent minors]\label{defn_Minor}
 For any $ (p+n)\times (p+n)$ matrix $\cal A$ and $\fa \in \mathcal I$, we define the minor $\cal A^{(\mathfrak c)}:=(\cal A_{\fa\fb}:\fa,\fb \in \mathcal I\setminus \{\mathfrak a\})$ as the $ (p+n-1)\times (p+n-1)$ matrix obtained by removing the $\mathfrak c$-th row and column in $\cal A$. Note that we keep the names of indices when defining $\cal A^{(\mathfrak c)}$, i.e. $(\cal A^{(\mathfrak c)})_{\fa\fb}= \cal A_{\fa\fb}$ for $\fa,\fb \ne \mathfrak c$. Correspondingly, we define the resolvent minor for $G$ in \eqref{resolv Gauss1} as %(recall equation \eqref{green2})
\begin{align*}
G^{(\mathfrak c)}:&=\left[ \left( {\begin{array}{*{20}c}
   { -z\id_{p} } & n^{-1/2}\Lambda Z_1^\top & n^{-1/2} Z_2^\top  \\
   {n^{-1/2} Z_1 \Lambda  } & {-\id_{n_1}} & 0 \\
   {n^{-1/2} Z_2} & 0 & {-\id_{n_2}}
   \end{array}} \right)^{(\mathfrak c)}\right]^{-1} ,
%= \left( {\begin{array}{*{20}c}
%   { \mathcal G^{(\mathbb T)}} & \mathcal G^{(\mathbb T)} W^{(\mathbb T)}  \\
%   {\left(W^{(\mathbb T)}\right)^\top\mathcal G^{(\mathbb T)}} & { \mathcal G_R^{(\mathbb T)} }  \\
%\end{array}} \right)  ,
\end{align*}
and define the partial traces $m^{(\mathfrak c)}$ and $m^{(\mathfrak c)}_k$, $k=0,1,2,$ by replacing $G$ with $G^{(\mathfrak c)}$ in equation \eqref{defm}. For convenience, we adopt the convention that for the resolvent minor $G^{(\mathfrak c)}$ defined as above, $G^{(\mathfrak c)}_{\fa\fb} = 0$ if $\fa =\mathfrak c$ or $\fb =\mathfrak c$.
%Note that the resolvent minor $G^{(\mathfrak c)}$ is defined such that it is independent of the entries in the $\mathfrak c$-th row and column of $H$. One will see a crucial use of this fact in the heuristic proof below.
 Using Schur complement formula (see {equation} (\ref{resolvent2})), we have that %for $i \in \cal I_0$, $\mu \in \cal I_1$ and $\nu\in \cal I_2$,
\HZ{This part needs more explanation - cannot understand.}
\begin{align}
\frac{1}{{G_{ii} }}&=  - z - \frac{\lambda_i^2}{n} \sum_{\mu,\nu\in \mathcal I_1} z_{\mu i}z_{\nu i}G^{\left( i \right)}_{\mu\nu} - \frac{1}{n} \sum_{\mu,\nu\in \mathcal I_2} z_{\mu i}z_{\nu i}G^{\left( i \right)}_{\mu\nu} -2 \frac{\lambda_i}{n} \sum_{\mu\in \cal I_1,\nu\in \mathcal I_2} z_{\mu i}z_{\nu i}G^{\left( i \right)}_{\mu\nu},  \ \ \text{for }\  i \in \cal I_0 , \label{0self_Gii}\\
\frac{1}{{G_{\mu\mu} }}&=  - 1 - \frac{1}{n} \sum_{i,j\in \mathcal I_0}\lambda_i \lambda_j z_{\mu i}z_{\mu j} G^{\left(\mu\right)}_{ij}, \quad \frac{1}{{G_{\nu\nu} }}=  - 1 - \frac{1}{n} \sum_{i,j\in \mathcal I_0}  z_{\nu i}z_{\nu j}  G^{\left(\nu\right)}_{ij},  \ \ \text{for }\  \mu \in \cal I_1, \ \nu\in \cal I_2, \label{0self_Gmu1}
\end{align}
where we recall the notations in equation \eqref{labelZ}. For the right-hand side of equation \eqref{0self_Gii}, notice that the resolvent minor $G^{(i)}$ is defined such that it is independent of the entries $z_{\mu i}$ and $z_{\nu i}$. Hence by the concentration inequalities in Lemma \ref{largedeviation}, we have that the  right-hand side of equation \eqref{0self_Gii} concentrates around the partial expectation over the entries $\{z_{\mu i}: \mu \in \cal I_1\cup \cal I_2\}$, i.e., with high probability,
\begin{align*}
\frac{1}{{G_{ii} }}&=  - z - \frac{\lambda_i^2}{n} \sum_{\mu \in \mathcal I_1}  G^{\left( i \right)}_{\mu\mu} - \frac{1}{n} \sum_{\mu\in \mathcal I_2} G^{\left( i \right)}_{\mu\mu} +\oo(1)= - z - \lambda_i^2 \frac{\rho_1}{\rho_1+\rho_2} m_1^{(i)}(z)-  \frac{\rho_2}{\rho_1+\rho_2} m_2^{(i)}(z)+\oo(1),
\end{align*}
where we used the definition of $m_1^{(i)}$ and $m_2^{(i)}$ in equation \eqref{defm} with $G$ replaced by $G^{(i)}$. Intuitively, since we have removed only one column and one row out of the $(p+n)$ columns and rows in $H$, $m_1^{(i)}$ and $m_2^{(i)}$ should be close to the original $m_1$ and $m_2$. Hence we obtain from the above equation that
\begin{align}\label{1self_Gii}
 G_{ii}  = -\left( z +\lambda_i^2 \frac{\rho_1}{\rho_1+\rho_2} m_1(z) +  \frac{\rho_2}{\rho_1+\rho_2}m_2(z)+\oo(1)\right)^{-1}.
\end{align}
Similarly, we can obtain from equation \eqref{0self_Gmu1} that for $\mu \in \cal I_1$ and $\nu\in \cal I_2$,
\be\label{1self_Gmu} G_{\mu \mu }=-\left(1+\frac{p}{n} m_0 + \oo(1)\right)^{-1},\quad G_{\nu\nu}=-\left(1+\frac{p}{n} m+\oo(1)\right)^{-1},\ee
with high probability.
\noindent\textbf{Deriving the self-consistent equations.} Taking average of \eqref{1self_Gmu}, we obtain that
\be\label{2self_Gmu} m_1= \frac{1}{n_1}\sum_{\mu \in \cal I_1}G_{\mu\mu}=-\left(1+\frac{p}{n} m_0 + \oo(1)\right)^{-1},\quad m_2=\frac{1}{n_2}\sum_{\nu \in \cal I_2}G_{\nu\nu}=-\left(1+\frac{p}{n} m+\oo(1)\right)^{-1},
with high probability. Together with the definition of $m$ and $m_0$ in equation \eqref{defm}, the two equations in equation \eqref{2self_Gmu} give that
\be\label{3self_Gmu}  \frac1{m_1}= -1- \frac{1}{n} \sum_{i=1}^p \lambda_i^2 G_{ii}+ \oo(1),\quad \frac1{m_2}=-1-\frac{1}{n} \sum_{i=1}^p G_{ii}  + \oo(1),\ee
with high probability. Plugging equation \eqref{1self_Gii} into equation \eqref{3self_Gmu}, we obtain that
\begin{align*}
& \frac1{m_1}= -1+ \frac{1}{n} \sum_{i=1}^p \frac{\lambda_i^2 }{ z +\lambda_i^2 \frac{\rho_1}{\rho_1+\rho_2} m_1(z) +  \frac{\rho_2}{\rho_1+\rho_2}m_2(z)+\oo(1)}+ \oo(1),\\
& \frac1{m_2}=-1+\frac{1}{n} \sum_{i=1}^p \frac{1}{ z +\lambda_i^2 \frac{\rho_1}{\rho_1+\rho_2} m_1(z) +  \frac{\rho_2}{\rho_1+\rho_2}m_2(z)+\oo(1)}  + \oo(1),
\end{align*}
with high probability, which give a system of approximate self-consistent equations for $(m_1,m_2)$. Compare them to the deterministic self-consistent equations in equation \eqref{selfomega_a}, one can observe that we should have
\be\label{approx m12 add}

>>>>

%\subsection{The Bias and Variance Asymptotics}
\paragraph{Proof overview.} 
%\todo{merge} Finally, in this subsection, we describe briefly the main ideas in the proofs of Lemmas \ref{lem_cov_shift} and \ref{lem_cov_derivative}.
%We first describe the proof of Theorem \ref{lem_cov_shift_informal}.
%We briefly describe the key ideas for proving Lemma \ref{lem_cov_shift} and Lemma \ref{lem_cov_derivative}.
We use the standard Stieltjes transform method (or the resolvent method) in random matrix theory \cite{bai2009spectral,tao2012topics,erdos2017dynamical}. Both Lemma \ref{lem_cov_shift} and Lemma \ref{lem_cov_derivative} require a study of the matrix inverse $[(X^{(1)})^\top X^{(1)}+(X^{(2)})^\top X^{(2)}]^{-1}$ for $X^{(1)}= Z^{(1)}\Sigma_1^{1/2}$ and $X^{(2)}= Z^{(2)}\Sigma_2^{1/2}$.
%\HZ{Suggestion: focus on the proof overview of Lemma \ref{lem_cov_shift} only. Use one paragraph in the end to say how to extend the idea to Lemma \ref{lem_cov_derivative}}
Suppose $M=\Sig_1^{1/2} \Sig_2^{-1/2}$ has a singular value decomposition
\be\label{eigen2}
M= U\Lambda V^\top, \quad \Lambda=\text{diag}( \lambda_1, \ldots, \lambda_p).
\ee
Then using equation \eqref{eigen2}, we can write
\be\label{eigen2extra}[(X^{(1)})^\top (X^{(1)})+(X^{(2)})^\top (X^{(2)})]^{-1}= \Sigma_2^{-1/2}V\left(   \Lambda U^\top (Z^{(1)})^\top Z^{(1)} U\Lambda  + V^\top (Z^{(2)})^\top Z^{(2)}V\right)^{-1}V^\top\Sigma_2^{-1/2}.\ee
We divide the discussion into the following parts.

@@

\noindent\textbf{Defining the resolvents.} For our purpose, we use a convenient linearization trick in linear algebra, that is, the SVD of a rectangular matrix $A$ is equivalent to the study of the eigendecomposition of the symmetric block matrix 
$$H(A):=\begin{pmatrix}0 & A \\ A^\top & 0\end{pmatrix},$$
which is linear in $A$. This trick has been used in many random matrix literatures, such as \cite{Anisotropic, AEK_Gram, XYY_circular,DY20201}. Our goal is to study the SVD of the rectangular matrix $(\Lambda U^\top (Z^{(1)})^\top,V^\top (Z^{(2)})^\top)$. %For this purpose, 
%}\HZ{what does this trick mean? use less technical words},

@@

Then we define the following $(p+n)\times (p+n)$ symmetric block matrix 
%which is a linear function of $(Z^{(1)})$ and $(Z^{(2)})$:

@@

    H\left(\Lambda U^\top (Z^{(1)})^\top,V^\top (Z^{(2)})^\top\right): = n^{-1/2}\left( {\begin{array}{*{20}c}
   { 0 } & \Lambda U^{\top}(Z^{(1)})^\top & V^\top (Z^{(2)})^\top  \\
   {Z^{(1)} U\Lambda  } & {0} & 0 \\
   {Z^{(2)}V} & 0 & 0

@@

For simplicity of notations, we abbreviate it as $H$ in the following discussion, and define the index sets

@@

 We will consistently use the latin letters $i,j\in\sI_{0}$ and greek letters $\mu,\nu\in\sI_{1}\cup \sI_{2}$. Correspondingly, the indices of the matrices $Z^{(1)}$ and $Z^{(2)}$ are labelled as

@@

 Z^{(1)}= (Z^{(1)}_{\mu i}:i\in \mathcal I_0, \mu \in \mathcal I_1), \quad Z^{(2)}= (Z^{(2)}_{\nu i}:i\in \mathcal I_0, \nu \in \mathcal I_2).\ee

@@

 G (z):= \left[H -\left( {\begin{array}{*{20}c}

@@

\end{array}} \right)\right]^{-1} , \quad z\in \mathbb C ,

@@

and the resolvent of $  n^{-1}\left(\Lambda U^\top (Z^{(1)})^\top Z^{(1)} U\Lambda  + V^\top (Z^{(2)})^\top Z^{(2)}V\right)$ as

@@

\cal G(z):=\left( n^{-1}  \Lambda U^\top (Z^{(1)})^\top Z^{(1)} U\Lambda  + n^{-1}V^\top (Z^{(2)})^\top Z^{(2)}V -z\right)^{-1},\quad z\in \C.

@@

We will show that these averaged resolvents satisfy some deterministic self-consistent equations asymptotically, which will be the core part of the proof. We refer the reader to the discussions below \eqref{0self_Gii} and \eqref{0self_Gmu1}.

@@

With Schur complement formula, we can check that %\HZ{Has $G(z)$ been defined before?}

@@

   { \cal G (z)} & \cal G(z)A  \\
   A^\top \cal G(z) & z (A^\top A - z)^{-1}

@@

 where we abbreviated $A:= n^{-1/2}(\Lambda U^\top (Z^{(1)})^\top, V^\top (Z^{(2)})^\top)$. 
 %\HZ{$W$ is overloaded with the output layers} 
 This shows that a control of $G(z)$ yields directly a control of $\mathcal G(z)$ as the upper-left block. On the other hand, $G(z)$ is a little  easier to use than $\cal G(z)$, and obviously contains more information.

@@

\medskip
\noindent\textbf{Defining the asymptotic limits of the resolvents.}
%\label{sec aymp_limit_G}
%\subsection{Resolvents and limiting law}
We now describe the asymptotic limit of $G(z)$ as $n\to \infty$. First we define the deterministic limits of $(m_1(z), m_{2}(z))$ by $\left(-\frac{\rho_1+\rho_2}{\rho_1}a_{1}(z),-\frac{\rho_1+\rho_2}{\rho_2}a_{2}(z)\right)$, where $(a_1(z), a_2(z))$ is
%\HZ{This $M_{1}, M_{2}$ notation is awkward; consider changing it to ${M}_1(z), {M}_2(z)$ (or  better, say $a_1(z), a_2(z)$? if they correspond to $a_1, a_2$ when $z=0$)},
the unique solution to the following system of equations
\begin{equation}\label{selfomega_a}
\begin{split}
& \frac{\rho_1}{a_{1}(z)} = \frac{1}{p}\sum_{i=1}^p \frac{\lambda_i^2}{ - z+\lambda_i^2 a_{1}(z) +a_{2} (z) } + (\rho_1+\rho_2),\  \frac{\rho_2}{a_{2}(z)} = \frac{1}{p}\sum_{i=1}^p \frac{1 }{  -z+\lambda_i^2 a_{1}(z) +  a_{2}(z)  }+ (\rho_1+\rho_2) ,
\end{split}
\ee
satisfying that $\im a_{1}(z)>0$ and $\im a_{2}(z)>0$ for $z\in \C_+$ with $\im z$.
%Here, for simplicity of notations, we introduced the following ratios
%\be\label{ratios}
% \gamma_n :=\frac{p}{n}=\frac{1}{\rho_1+\rho_2},\quad r_1 :=\frac{n_1}{n}=\frac{\rho_1}{\rho_1+\rho_2},\quad r_2 :=\frac{n_2}{n}=\frac{\rho_2}{\rho_1+\rho_2}.
%\ee
We define the matrix limit of $G(z)$ as
\be \label{defn_piw}
\Gi(z) := \begin{pmatrix} (-z\id_p+a_{1}(z)\Lambda^2  +  a_{2}(z)\id_p)^{-1} & 0 & 0 \\ 0 & - \frac{\rho_1+\rho_2}{\rho_1} a_{1}(z)\id_{n_1} & 0 \\ 0 & 0 & -\frac{\rho_1+\rho_2}{\rho_2}a_{2}(z)\id_{n_2}  \end{pmatrix}.\ee
In particular, the matrix limit of $\cal G(z)$ is given by 
\be\label{matrix limit}(-z\id_p+a_{1}(z)\Lambda^2 + a_{2}(z)\id_p)^{-1}.\ee

\medskip
\noindent{\bf Deriving the variance asymptotics.} Using definition \eqref{mainG}, we can write equation \eqref{eigen2extra} as
\be\label{rewrite X as R} [(X^{(1)})^\top X^{(1)}+(X^{(2)})^\top X^{(2)}]^{-1}=n^{-1}\Sigma_2^{-1/2}V\cal G(0)V^\top\Sigma_2^{-1/2}.\ee
When $z=0$, it is easy to check that equation \eqref{selfomega_a} can be reduced to \eqref{eq_a12extra}, which means that we actually have $a_1(0)=a_1$ and $a_2(0)=a_2$. Hence the matrix limit of $\cal G(0)$ is given by $(a_{1}\Lambda^2 + a_{2}\id_p)^{-1}$. Then inserting this limit into \eqref{rewrite X as R}, we can write the left-hand side of \eqref{lem_cov_shift_eq} as
\begin{align*}
&\bigtr{(X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} \Sigma}\approx n^{-1}\bigtr{\Sigma_2^{-1/2}V\cal (a_{1}\Lambda^2 + a_{2}\id_p)^{-1}V^\top\Sigma_2^{-1/2}\Sigma}  \\
&=n^{-1}\bigtr{\Sigma_2^{-1/2}\cal (a_{1}\Sigma_2^{-1/2}\Sigma_1\Sigma_2^{-1/2} + a_{2}\id_p)^{-1}\Sigma_2^{-1/2}\Sigma}  = n^{-1}\bigtr{\cal (a_{1} \Sigma_1  + a_{2}\Sigma_2)^{-1}\Sigma}  ,
\end{align*}
where in the second step we used $V \Lambda^2 V^\top=M^\top M=\Sigma_2^{-1/2}\Sigma_1\Sigma_2^{-1/2}$ by equation \eqref{eigen2}.  This concludes \eqref{lem_cov_shift_eq} up to a small error.

%\alert{Together with equation \eqref{rewrite X as R} and $V \Lambda^2 V^\top=M^\top M=\Sigma_2^{-1/2}\Sigma_1\Sigma_2^{-1/2}$ by equation \eqref{eigen2}, we can obtain equation \eqref{lem_cov_shift_eq} in the large $n$ limit.}\HZ{I do not understand this sentence.}

\medskip
\noindent{\bf Gaussian case.} Now we give a heuristic derivation of the matrix limit when the entries of $Z^{(1)}$ and $Z^{(2)}$ are i.i.d. Gaussian. In this case,
%However, notice that if the entries of $(Z^{(1)})\equiv (Z^{(1)})^{\text{Gauss}}$ and $(Z^{(2)})\equiv (Z^{(2)})^{\text{Gauss}}$ are i.i.d. Gaussian, then
by the rotational invariance of the multivariate Gaussian distribution we have
\be\label{eq in Gauss} Z^{(1)} U\Lambda \stackrel{d}{=} Z^{(1)} \Lambda, \quad Z^{(2)} V \stackrel{d}{=} Z^{(2)},\ee
where ``$\stackrel{d}{=}$" means ``equal in distribution". Hence it suffices to consider the following resolvent
 \begin{equation} \label{resolv Gauss1}
   G(z)= \left( {\begin{array}{*{20}c}
   { -z\id_{p} } & n^{-1/2}\Lambda (Z^{(1)})^\top & n^{-1/2} (Z^{(2)})^\top  \\
   {n^{-1/2} Z^{(1)} \Lambda  } & {-\id_{n_1}} & 0 \\
   {n^{-1/2} Z^{(2)}} & 0 & {-\id_{n_2}}
   \end{array}} \right)^{-1}.
 \end{equation}

\medskip
\noindent{\bf Schur complements.}
%Now we discuss about how to obtain the matrix limit in equation \eqref{defn_piw}.
%\HZ{Divide into several parts like sec 7 to improve readability.}
The core quantities of the derivation are the following resolvent minors, which are defined by removing certain rows and columns of the matrix $H$.
\begin{definition}[Resolvent minors]\label{defn_Minor}
 For any $ (p+n)\times (p+n)$ matrix $\cal A$ and $a\in \mathcal I$, we define the minor $\cal A^{(a)}:=(\cal A_{a_1 a_2}:a_1, a_2 \in \mathcal I\setminus \{a\})$ as the $ (p+n-1)\times (p+n-1)$ matrix obtained by removing the $a$-th row and column in $\cal A$. Note that we keep the names of indices when defining $\cal A^{(a)}$, i.e. $(\cal A^{(a)})_{a_1a_2}= \cal A_{a_1 a_2}$ for $a_1,a_2\ne a$. Correspondingly, we define the resolvent minor for $G$ in \eqref{resolv Gauss1} as %(recall equation \eqref{green2})
\begin{align*}
G^{(\mathfrak c)}:&=\left[ \left( {\begin{array}{*{20}c}
   { -z\id_{p} } & n^{-1/2}\Lambda (Z^{(1)})^\top & n^{-1/2} (Z^{(2)})^\top  \\
   {n^{-1/2} Z^{(1)} \Lambda  } & {-\id_{n_1}} & 0 \\
   {n^{-1/2} Z^{(2)}} & 0 & {-\id_{n_2}}
   \end{array}} \right)^{(a)}\right]^{-1} ,
%= \left( {\begin{array}{*{20}c}
%   { \mathcal G^{(\mathbb T)}} & \mathcal G^{(\mathbb T)} W^{(\mathbb T)}  \\
%   {\left(W^{(\mathbb T)}\right)^\top\mathcal G^{(\mathbb T)}} & { \mathcal G_R^{(\mathbb T)} }  \\
%\end{array}} \right)  ,
\end{align*}
and define the partial traces $m^{(a)}$, $m_0^{(a)}$, $m_1^{(a)}$ and $m_2^{(a)}$ by replacing $G$ with $G^{(a)}$ in equation \eqref{defm}. We adopt the convention that for the resolvent minor $G^{(a)}$ defined as above, $G^{(a)}_{a_1a_2} = 0$ if $a_1 =a$ or $a_2=a$.
\end{definition}
%Note that the resolvent minor $G^{(\mathfrak c)}$ is defined such that it is independent of the entries in the $\mathfrak c$-th row and column of $H$. One will see a crucial use of this fact in the heuristic proof below.
 Using Schur complement formula (see {equation} (\ref{resolvent2})), we have that for $i \in \cal I_0$, %$\mu \in \cal I_1$ and $\nu\in \cal I_2$,
%\HZ{This part needs more explanation - cannot understand.}
\begin{align}
\frac{1}{{G_{ii} }}&=  - z - \frac{\lambda_i^2}{n} \sum_{\mu,\nu\in \mathcal I_1} Z^{(1)}_{\mu i}Z^{(1)}_{\nu i}G^{\left( i \right)}_{\mu\nu} - \frac{1}{n} \sum_{\mu,\nu\in \mathcal I_2} Z^{(2)}_{\mu i}Z^{(2)}_{\nu i}G^{\left( i \right)}_{\mu\nu} -2 \frac{\lambda_i}{n} \sum_{\mu\in \cal I_1,\nu\in \mathcal I_2} Z^{(1)}_{\mu i}Z^{(2)}_{\nu i}G^{\left( i \right)}_{\mu\nu} , \label{0self_Gii}
\end{align}
and for $\mu \in \cal I_1$ and $\nu\in \cal I_2$,
\begin{align}
\frac{1}{{G_{\mu\mu} }}&=  - 1 - \frac{1}{n} \sum_{i,j\in \mathcal I_0}\lambda_i \lambda_j Z^{(1)}_{\mu i}Z^{(1)}_{\mu j} G^{\left(\mu\right)}_{ij}, \quad \frac{1}{{G_{\nu\nu} }}=  - 1 - \frac{1}{n} \sum_{i,j\in \mathcal I_0}  Z^{(2)}_{\nu i}Z^{(2)}_{\nu j}  G^{\left(\nu\right)}_{ij},
\label{0self_Gmu1}
\end{align}
where we recall the notations in equation \eqref{labelZ}. To see why equation \eqref{0self_Gii} holds, we have that by Schur complement formula,
\begin{align*}
\frac{1}{G_{ii}}= -z - H_i G^{(i)}H_{i}^\top, \quad H_i = \left( \mathbf 0_{p-1}, (n^{-1/2}\lambda_i (Z^{(1)})^\top_{i\mu }:\mu \in \cal I_1),(n^{-1/2} (Z^{(2)})^\top_{i\nu }:\nu \in \cal I_2)\right),
\end{align*}
where $H_i$ is actually the $i$-th row of $H$ with the $(i,i)$-th entry removed. Expanding the above expression, we obtain equation \eqref{0self_Gii}. The two expressions in \eqref{0self_Gmu1} are easier to obtain using Schur complement formula.
